Extended Data

### 2021-10-18 doc/manuscript/Dow_manuscript_Nature.Rmd
fill in references, misc cleanup

As climate change accelerates and spring temperatures become increasingly warmer, growing seasons will start earlier; however, barring rapid acclimation of forests to the warming conditions, this is unlikely to provide the sustained increase in carbon sequestration and ensuant negative climate change feedback that is anticipated in most models (Fig. `r schematic_summary`)[@buermann_widespread_2018; @zani_increased_2020;@keenan_net_2014].
Rather, the dominant effect of rising temperatures on forest woody productivity will be a negative effect of high summer temperatures (Fig. `r tree_rings`), which constitutes a positive feedback to climate change.

@@

*We have not yet re-written the methods for Nature. Nature guidelines: Full methods, typically not exceeding 3000 words (currently at 1800 words). May be subdivided by bold headings.*

@@

 <!--Climate -->
Climate data corresponding to the measurement periods at both sites were obtained from local weather stations. 
For SCBI, weather data were obtained from a meteorological tower adjacent to the SCBI ForestGEO plot, via the ForestGEO Climate Data Portal v1.0 (https://forestgeo.github.io/Climate/)[@anderson-teixeira_forestgeo_2020]. 
Daily temperature records for Harvard Forest were obtained from the Harvard Forest weather station **(Boose & Gould, 2004; Boose, 2018)** ([GitHub issue #81](https://github.com/EcoClimLab/growth_phenology/issues/81)).
The R package *clim.pact*[@benestad_empiricalstatistical_2008] was used to plot temperatures for visual inspection and to identify readings that were >3 standard deviations away from yearly means, which were labeled as outliers and removed from the dataset. 
Gaps in the SCBI meteorological tower data were subsequently filled using temperature readings obtained from a National Center for Environmental Information (NCEI) weather station located in Front Royal, Virginia (https://www.ncdc.noaa.gov/cdo-web/datasets/GHCND/stations/GHCND:USC00443229/detail). 
For each site, we extracted daily maximum ($T_{max}$) and minimum temperatures ($T_{min}$).

@@

Correlation between growth parameters ($DOY_{25}$, $DOY_{50}$, $DOY_{75}$, $DOY_{g_{max}}$, $L_{PGS}$, $g_{max}$, and $\Delta DBH$}, Fig. `r schematic_summary`) and spring temperatures were assessed using a linear mixed model in a hierarchical Bayesian framework.
Analyses were run for both $T_{max}$ and $T_{min}$, with qualitatively similar results, but we present only results for $T_{max}$, which had overall stronger correlation with growth parameters. 
Mixed effects models were used to test the response of growth phenology variables to fixed effects of xylem porosity and mean $T_{max}$ (or $T_{min}$) during the critical temperature window selected by *climwin*, along with random effects of species and tree. 
We ran separate models for each species group at each site, and for the response of all growth phenology variables to $T_{max}$ (or $T_{min}$). 
Models were run using the LME4 package in R [@bates_fitting_2015]. 
This mixed-effect model was run within a hierarchical Bayesian framework and fit using the `rstanarm` R interface to the Stan programming language @rstanarm_2020. In all cases unless otherwise specified, all prior distributions are set to be the weakly informative defaults. 
<!--
TODO added by Bert: Whereever we specify model details, include details about prior from here:
https://cran.r-project.org/web/packages/rstanarm/vignettes/priors.html#default-weakly-informative-prior-distributions
-->

@@

We obtained monthly $T_{max}$ and $T_{min}$ data for 1901-2019 from CRU v.4.04 [@harris_version_2020] via the ForestGEO Climate Data Portal v1.0 [@anderson-teixeira_forestgeo_2020].

<!--QUILT PLOT METHODS-->
Correlations between monthly climate and $RWI$ were assessed using ‘dplR’[@bunn_dendrochronology_2008] and ‘bootRes’[@zang_dendroclimatic_2013] in R v.**##** (R Core Team, **YEAR**), which correlated functions and bootstrapped confidence intervals for these relationships[@biondi_dendroclim2002_2004].
We analyzed these correlations for January through September of the current year (presented in Fig. `r tree_rings`, `r EDF_tree_rings_TMIN`).
To test for potential lag effects effects of spring temperatures on the following year's growth, we also ran a version of the analysis extending back to include climate of every month of the previous year (`r EDF_tree_rings_TMAX_extended`).

<!--MULTIVARIATE ANALYSIS METHODS-->
**MULTIVARIATE ANALYSIS METHODS**

# Acknowledgements {-#Acknowledgements} 
people: (everyone who helped collect data-- see list on GitHub), Jess Shue (data)
Funding: ForestGEO

@@

>>>>

As climate change accelerates and spring temperatures become increasingly warmer, growing seasons will start earlier; however, barring rapid acclimation of forests to the warming conditions, this is unlikely to provide the sustained increase in carbon sequestration and ensuant negative climate change feedback that is anticipated in most models[@buermann_widespread_2018; @zani_increased_2020;@keenan_net_2014].
Rather, the dominant effect of rising temperatures on forest woody productivity will be a negative effect of high summer temperatures, which constitutes a positive feedback to climate change.

@@

<!--Nature guidelines: Full methods, typically not exceeding 3000 words. May be subdivided by bold headings.-->
<!--We are currently at ~2000 words -->

@@

 <!--Climate -->
Climate data corresponding to the measurement periods at both sites were obtained from local weather stations. 
For SCBI, weather data were obtained from a meteorological tower adjacent to the SCBI ForestGEO plot, via the ForestGEO Climate Data Portal v1.0 (https://forestgeo.github.io/Climate/)[@anderson-teixeira_forestgeo_2020]. 
Daily temperature records for Harvard Forest were obtained from the Harvard Forest weather station **(Boose & Gould, 2004; Boose, 2018)** (**CAM, we need these references. And see [GitHub issue #81](https://github.com/EcoClimLab/growth_phenology/issues/81)**).
The R package *clim.pact*[@climpact_REF; @benestad_empiricalstatistical_2008] was used to plot temperatures for visual inspection and to identify readings that were >3 standard deviations away from yearly means, which were labeled as outliers and removed from the dataset. 
Gaps in the SCBI meteorological tower data were subsequently filled using temperature readings obtained from a National Center for Environmental Information (NCEI) weather station located in Front Royal, Virginia (https://www.ncdc.noaa.gov/cdo-web/datasets/GHCND/stations/GHCND:USC00443229/detail). 
For each site, we extracted daily maximum ($T_{max}$) and minimum temperatures ($T_{min}$).

@@

Correlation between growth parameters ($DOY_{25}$, $DOY_{50}$, $DOY_{75}$, $DOY_{g_{max}}$, $L_{PGS}$, $g_{max}$, and $\Delta DBH$}, Fig. `r schematic_summary`) and spring temperatures were assessed using a linear mixed model in a hierarchical Bayesian framework.
Analyses were run for both $T_{max}$ and $T_{min}$, with qualitatively similar results, but we present only results for $T_{max}$, which had overall stronger correlation with growth parameters. 
Mixed effects models were used to test the response of growth phenology variables to fixed effects of xylem porosity and mean $T_{max}$ (or $T_{min}$) during the critical temperature window selected by *climwin*, along with random effects of species and tree. 
We ran separate models for each species group at each site, and for the response of all growth phenology variables to $T_{max}$ (or $T_{min}$). 
Models were run using the LME4 package in R [@bates_fitting_2015]. 
This mixed-effect model was run within a hierarchical Bayesian framework and fit using the `rstanarm` R interface to the Stan programming language[@gabry_rstanarm_2020]. In all cases unless otherwise specified, all prior distributions are set to be the weakly informative defaults. 
<!--
TODO added by Bert: Whereever we specify model details, include details about prior from here:
https://cran.r-project.org/web/packages/rstanarm/vignettes/priors.html#default-weakly-informative-prior-distributions
-->

@@

We obtained monthly $T_{max}$ and $T_{min}$ data for 1901-2019 from CRU v.4.04 [@harris_version_2020].

<!--QUILT PLOT METHODS-->
Correlations between monthly climate and $RWI$ were assessed using ‘dplR’[@bunn_dendrochronology_2008] and ‘bootRes’[@zang_dendroclimatic_2013] in R v.**##** (R Core Team, **YEAR**), which correlated functions and bootstrapped confidence intervals for these relationships[@biondi_dendroclim2002_2004].
We analyzed these correlations for January through September of the current year (presented in Fig. `r tree_rings`, `r EDF_tree_rings_TMIN`).
To test for potential lag effects effects of spring temperatures on the following year's growth, we also ran a version of the analysis extending back to include climate of every month of the previous year (`r EDF_tree_rings_TMAX_extended`).
Correlations and significance levels for months April-August are given in **APPENDIX**.

<!--MULTIVARIATE ANALYSIS METHODS-->
We used a multivariate model to test for joint effects of April and summer $T_{max}$ on RWI.
We began by testing univariate correlations of $T_{max}$ over three summer windows: June, June-July, and May-August. 
Having determined that, among these, June-July explained the most variation, we then analyzed the joint effects of April $T_{max}$ and June-July $T_{max}$ on RWI for each chronology independently using **(Function)** in R. 
Slopes and p-values for each chronology are given in **APPENDIX**.

# Acknowledgements {-#Acknowledgements} 
- people: (everyone who helped collect data-- see list on GitHub), Jess Shue (data)

- Funding: ForestGEO

**COAUTHORS, PLEASE ADD.**

@@

**(CAM, please draft this before sending around.)**
